---
title: "Causal Inference Final Code"
author: "Sam Neylon"
date: '2024-05-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(stringr)
library(scales)

set.seed(7008)

```

# Import

```{r }

# Import ln_emp_flood_df

ln_emp_flood_df <- read_csv(here("data/SHELDUS/ln_emp_flood_df.csv"),
                     col_types = cols(
  year = col_double(),
  st = col_character(),
  cty = col_character(),
  firms = col_double(),
  estabs = col_double(),
  emp = col_double(),
  denom = col_double(),
  estabs_entry = col_double(),
  estabs_entry_rate = col_double(),
  estabs_exit = col_double(),
  estabs_exit_rate = col_double(),
  job_creation = col_double(),
  job_creation_births = col_double(),
  job_creation_continuers = col_double(),
  job_creation_rate_births = col_double(),
  job_creation_rate = col_double(),
  job_destruction = col_double(),
  job_destruction_deaths = col_double(),
  job_destruction_continuers = col_double(),
  job_destruction_rate_deaths = col_double(),
  job_destruction_rate = col_double(),
  net_job_creation = col_double(),
  net_job_creation_rate = col_double(),
  reallocation_rate = col_double(),
  firmdeath_firms = col_double(),
  firmdeath_estabs = col_double(),
  firmdeath_emp = col_double(),
  FIPS_5 = col_character(),
  PropertyDmg = col_double(),
  PropertyDmg_Adj2022 = col_double(),
  PropertyDmgPerCapita_Adj2022 = col_double(),
  SHELDUS = col_double(),
  pre_96 = col_double(),
  serious_flood = col_double(),
  treatment = col_double(),
  first_hurr_year = col_double(),
  group = col_double(),
  ln_emp = col_double(),
  time_to_treat = col_double(),
  treat = col_double(),
  FIPS = col_double(),
  treatment_post = col_double()
))

```

# Treatment Variables

I prepared these different treatment variables (different packages required the treatment variables to be in different forms) in other code files, but I am reproducing them here for reference.

## Serious Flood Threshold

Setting the flooding damage threshold at 95th percentile for now (based on some analysis I did on ChatGPT).

```{r eval=FALSE}

serious_threshold <- floodSHELDUS_bds_df %>%
  filter(SHELDUS == 1) %>%  
  summarise(quantile_95 = quantile(PropertyDmgPerCapita_Adj2022, 0.95, na.rm = TRUE)) %>%
  pull(quantile_95) 

# Check the calculated threshold
serious_threshold

```

## 'treatment' variable

'treatment' is an indicator for the post-treatment period for each treated county.

```{r eval=FALSE}

floodSHELDUS_bds_df <- floodSHELDUS_bds_df %>%
  # Define a serious flood
  mutate(serious_flood = if_else(PropertyDmgPerCapita_Adj2022 >= serious_threshold, 1, 0)) %>%
  # Group by county
  group_by(FIPS_5) %>%
  # Create the treatment variable
  mutate(treatment = cummax(serious_flood)) %>%
  ungroup()

```

## 'group' variable

Used for Callaway and Sant'Anna (2021) Group-Time ATT, which I didn't end up putting in the final paper, but I use the 'group' variable to build other treatment variables.

'group' is the first year a county has a serious flood, and is 0 if never-treated.

```{r eval=FALSE}

floodSHELDUS_bds_df <- floodSHELDUS_bds_df %>%
  arrange(FIPS_5, year) %>%
  group_by(FIPS_5) %>%
  mutate(
    first_hurr_year = ifelse(any(serious_flood == 1), min(year[serious_flood == 1]), NA_real_),
    group = ifelse(is.na(first_hurr_year), 0, first_hurr_year)
  ) %>%
  ungroup()

```

## 'time_to_treat' & 'treat'

Variables for 'fixest' package.

'time_to_treat' is the event study, with -1000 for never-treated counties (following 'fixest' documentation).

'treat' is a dummy for whether a county is ever treated.

```{r eval=FALSE}

flood_df <- flood_df %>% 
  mutate(
    time_to_treat = ifelse(group > 0, year - group, -1000),
    treat = ifelse(group > 0, 1, 0)
  )

```

## 'treatment_post'

'treatment_post' is an interaction of 'treatment' (the post-treatment period indicator) and 'treat' (the ever-treated indicator). The coefficient of this interaction is the traditional difference-in-differences ATT estimate.

```{r eval=FALSE}

ln_emp_flood_df$treatment_post <- ln_emp_flood_df$treatment * ln_emp_flood_df$treat

```

# Descriptives

## Continuous

```{r}

cont_desc_fn <- function(data, variables) {
  stats <- data %>%
    summarise(across(all_of(variables), 
                    list(N = ~sum(!is.na(.)),
                          Mean = ~round(mean(., na.rm = TRUE), 2),
                          P25 = ~round(quantile(., 0.25, na.rm = TRUE), 2),
                          P50 = ~round(quantile(., 0.50, na.rm = TRUE), 2),
                          P75 = ~round(quantile(., 0.75, na.rm = TRUE), 2),
                          P95 = ~round(quantile(., 0.95, na.rm = TRUE), 2),
                          Max = ~round(max(., na.rm = TRUE), 2)),
                     .names = "{.col}__{.fn}"))
  
  stats_long <- stats %>%
    pivot_longer(cols = everything(), names_to = c("Variable", "Statistic"), 
                 names_sep = "__") %>%
    pivot_wider(names_from = Statistic, values_from = value)
  
  return(stats_long)
}

# Example usage
# Assuming df is your data frame and you have a list of variables you want to summarize
cont_vars <- c("emp", "ln_emp")
cont_vars_noZero <- c("PropertyDmg_Adj2022", "PropertyDmgPerCapita_Adj2022")

cont_tbl_nZ <- ln_emp_flood_df %>% 
  filter(PropertyDmg_Adj2022 > 0) %>% 
  cont_desc_fn(data = ., variables = cont_vars_noZero)

cont_tbl <- cont_desc_fn(ln_emp_flood_df, cont_vars)

cont_tbl_full <- bind_rows(cont_tbl, cont_tbl_nZ)

cont_tbl_full <- cont_tbl_full %>%
  mutate(across(-c(Variable, N), ~case_when(
    Variable == "emp" ~ paste0(comma(. / 1000, accuracy = 0.01), "k"),
    Variable == "ln_emp" ~ paste0(comma(., accuracy = 0.01)),
    Variable == "PropertyDmg_Adj2022" ~ paste0("$", comma(. / 1000, accuracy = 0.01), "k"),
    Variable == "PropertyDmgPerCapita_Adj2022" ~ paste0("$", comma(., accuracy = 0.01)))
  ))

```

```{r}

# Print the summary table
print(cont_tbl_full)

```

## Avg. Floods per Year

```{r eval=FALSE}

avg_floods_per_year <- ln_emp_flood_df %>%
  filter(group != 0) %>%
  distinct(FIPS, .keep_all = TRUE) %>%
  group_by(group) %>%
  summarize(counties_hit = n()) %>%
  summarize(average_per_year = mean(counties_hit))

# Print the result
print(avg_floods_per_year)

```

# Vanilla OLS

```{r }

ols.ln_emp.model <- lm(ln_emp ~ treatment_post, data = ln_emp_flood_df)

```

```{r }

summary(ols.ln_emp.model)

```

# fixest

```{r }

library(fixest)

```

## Model 1 - Averaged ATT

```{r }

feols_ln_emp.AvgATT <- feols(ln_emp ~ treatment_post | FIPS_5 + year, ln_emp_flood_df, cluster = "FIPS_5")

```

```{r }

summary(feols_ln_emp.AvgATT)

```

## Event Study - Time-varying ATT's

```{r }

feols_ln_emp.Event <- feols(ln_emp ~ i(time_to_treat, treat, ref = c(-1, -1000)) | FIPS_5 + year, ln_emp_flood_df, cluster = "FIPS_5")

```

### Summary

```{r }

summary(feols_ln_emp.Event)

```

### Figure X - Event Plot

```{r }

iplot(feols_ln_emp.Event)

```

# fect

```{r }

library(fect)
library(panelView)

```

## Model

```{r }

fect_ln_emp <- fect(ln_emp ~ treatment, 
                data = ln_emp_flood_df, 
                index = c("FIPS","year"),
                # force: include both unit and time effects
                force = "two-way",
                # Interactive Fixed Effects
                method = "ife", 
                # Cross-validation to pick number of factors
                CV = TRUE, 
                # Test between 0 and 5 factors
                r = c(0, 5), 
                se = TRUE, 
                nboots = 200, 
                parallel = TRUE,
                #cores = 4) # 4 for GC pc
                cores = 6) # 6 for home PC

```

## Summary

```{r }

print(fect_ln_emp)

```

### Plot

```{r }

plot(fect_ln_emp, main = "Estimated ATT (IFEct)")

```

```{r}

fect_ln_emp$est.att

```

