---
title: "Flood-BDS Analysis 1-3-24"
output: html_document
date: "2024-01-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(stringr)
# library(scales)

set.seed(7008)

```

# Notes

##(1-2-24)

Lots of ChatGPT driven flood threshold analysis.

NOTE: I ran all the analysis on a version of the data which eliminated zeroes.

##(1-4-24)

I have created a series of copies of the "flood-BDS_analysis" file, in order to try out different things.

NOTE: As the file title says (at the end) this version says that "serious floods" are defined as being at the 95th percentile of flooding damage per employee.

This file is specifically for exploring IFECT models.

I don't know how IFECT handles imbalanced panels - I think it might try to work with them, instead of deleting them, so I need to make sure and delete counties with missing 'emp' values.

THERE WAS A MAJOR PROBLEM! I was accidentally creating a bunch of NA's in my group variable. Unfortunately, I've found that ChatGPT loves to use 'any()' in if_else statements, and doesn't take into account the vector lengths - something to be aware of in the future!

I have now:
  - Renamed the files with bad 'group' variables "BROKE_"
  - Am now creating new versions, with fixed group.
    - fect is done
    - 95 threshold is done
    - NOT DONE: 99 threshold (one with data "_1-3-24")

# Import

```{r}

flood_df <- read_csv(here("data/merge/flood-BDS_data.csv"),
                     col_types = cols(
  year = col_double(),
  st = col_character(),
  cty = col_character(),
  firms = col_double(),
  estabs = col_double(),
  emp = col_double(),
  denom = col_double(),
  estabs_entry = col_double(),
  estabs_entry_rate = col_double(),
  estabs_exit = col_double(),
  estabs_exit_rate = col_double(),
  job_creation = col_double(),
  job_creation_births = col_double(),
  job_creation_continuers = col_double(),
  job_creation_rate_births = col_double(),
  job_creation_rate = col_double(),
  job_destruction = col_double(),
  job_destruction_deaths = col_double(),
  job_destruction_continuers = col_double(),
  job_destruction_rate_deaths = col_double(),
  job_destruction_rate = col_double(),
  net_job_creation = col_double(),
  net_job_creation_rate = col_double(),
  reallocation_rate = col_double(),
  firmdeath_firms = col_double(),
  firmdeath_estabs = col_double(),
  firmdeath_emp = col_double(),
  FIPS_5 = col_character(),
  prop_dmg = col_double(),
  pre_96 = col_double()
                     ))

```

# Flooding Threshold

I want to take a data-driven approach to determining what counts as a serious flood.

I am using Damage per Employee as my metric, which may not be ideal, but it has been difficult to get county population data.

[Analysis suggested by ChatGPT!]

```{r eval=FALSE}

# Focus on non-zero data
flood_nonzero <- flood_df %>% filter(dmg_perEMP > 0)

```

## Histogram

```{r eval=FALSE}

# Assuming your data is in a dataframe named 'flood_data'
flood_nonzero %>%
  ggplot(aes(x = dmg_perEMP)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Distribution of Damage per Employee",
       x = "Damage per Employee",
       y = "Frequency")


```

## Stats Summary

```{r eval=FALSE}

flood_nonzero %>%
  summarise(Minimum = min(dmg_perEMP, na.rm = TRUE),
            Maximum = max(dmg_perEMP, na.rm = TRUE),
            Mean = mean(dmg_perEMP, na.rm = TRUE),
            Median = median(dmg_perEMP, na.rm = TRUE),
            SD = sd(dmg_perEMP, na.rm = TRUE))


```

## ChatGPT Visual

```{r eval=FALSE}

# Define a sequence of percentiles from 1st to 99th in 1% increments
percentile_seq <- seq(0.01, 0.99, by = 0.01)

# Function to calculate the number of unique counties above each percentile threshold
count_treated_counties <- function(percentile, data, value_column, fips_column) {
  threshold <- quantile(data[[value_column]], probs = percentile, na.rm = TRUE)
  data %>% 
    filter(!!sym(value_column) > threshold) %>% 
    summarise(count = n_distinct(!!sym(fips_column)))
}

# Applying the function across all percentiles
treated_counts <- map_df(percentile_seq, count_treated_counties, data = flood_nonzero, value_column = "dmg_perEMP", fips_column = "FIPS_5")

# Adding percentile information to the results
treated_counts <- treated_counts %>% 
  mutate(percentile = percentile_seq * 100)

# Plotting the results
ggplot(treated_counts, aes(x = percentile, y = count)) +
  geom_line() +
  geom_point() +
  labs(title = "Impact of Percentiles on the Number of Treated Counties",
       x = "Percentile of Damage per Employee",
       y = "Number of Treated Counties") +
  theme_minimal()


```

```{r eval=FALSE}

# Calculate the 90th percentile threshold
threshold_90th <- quantile(flood_nonzero$dmg_perEMP, 0.9, na.rm = TRUE)

# Filter the data for the top 10% (above the 90th percentile)
top_damage_df <- flood_nonzero %>%
  filter(dmg_perEMP > threshold_90th)

# Visualize the distribution of damage per employee for these top counties
ggplot(top_damage_df, aes(x = dmg_perEMP)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  scale_x_continuous(labels = comma) +  # Use comma formatting for the x-axis labels
  theme_minimal() +
  labs(title = "Distribution of Damage per Employee for Top 10% Counties",
       x = "Damage per Employee",
       y = "Frequency")



```
```{r eval=FALSE}

# Log transform and histogram
ggplot(top_damage_df, aes(x = dmg_perEMP)) +
  geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "blue", color = "black") +
  scale_x_log10(labels = scales::comma) +
  theme_minimal() +
  labs(title = "Log Transformed Distribution of Damage per Employee for Top 10% Counties",
       x = "Log of Damage per Employee",
       y = "Density")


```

```{r eval=FALSE}

ggplot(top_damage_df, aes(x = dmg_perEMP)) +
  geom_density(fill = "blue", alpha = 0.5) +
  scale_x_continuous(labels = scales::comma) +
  theme_minimal() +
  labs(title = "Density of Damage per Employee for Top 10% Counties",
       x = "Damage per Employee",
       y = "Density")


```

```{r eval=FALSE}

# Adjust the binwidth as needed
bin_width <- 1000  # Example bin width

ggplot(top_damage_df, aes(x = dmg_perEMP)) +
  geom_histogram(binwidth = bin_width, fill = "blue", color = "black") +
  scale_x_continuous(labels = scales::comma) +
  theme_minimal() +
  labs(title = "Histogram of Damage per Employee for Top 10% Counties",
       x = "Damage per Employee",
       y = "Frequency")


```

### 80th Pct

```{r eval=FALSE}

# Calculate the 80th percentile threshold
threshold_80th <- quantile(flood_nonzero$dmg_perEMP, 0.80, na.rm = TRUE)

# Filter the data for the top 10% (above the 90th percentile)
top20_damage_df <- flood_nonzero %>%
  filter(dmg_perEMP > threshold_80th)

ggplot(top20_damage_df, aes(x = dmg_perEMP)) +
  geom_histogram(aes(y = after_stat(density)), bins = 30, fill = "blue", color = "black") +
  scale_x_log10(labels = scales::comma) +
  theme_minimal() +
  labs(title = "Log Transformed Distribution of Damage per Employee for Top 20% Counties",
       x = "Log of Damage per Employee",
       y = "Density")



```

### Percentiles

```{r eval=FALSE}

# Calculate every 5th percentile for 'damage per employee'
percentiles <- seq(0, 1, by = 0.05)
percentile_values <- sapply(percentiles, function(p) quantile(flood_nonzero$dmg_perEMP, p, na.rm = TRUE))

# Modified logarithm function that adds a small value to handle zeros
log_mod <- function(x) log(x + 1e-6)

# Calculate the modified logarithm of the percentile values
log_percentile_values <- sapply(percentile_values, log_mod)

# Combine into a dataframe and format to avoid scientific notation
percentile_table <- data.frame(
  Percentile = percentiles,
  DamagePerEmployee = format(percentile_values, scientific = FALSE),
  LogDamagePerEmployee = format(log_percentile_values, scientific = FALSE)
)

# Print the table
print(percentile_table)

```

### Counties per Cutoff

```{r eval=FALSE}


# Define a function to count unique counties above a percentile threshold
count_treated_counties <- function(percentile, df, dmg_column, fips_column) {
  threshold <- quantile(df[[dmg_column]], probs = percentile, na.rm = TRUE)
  df %>% 
    filter(!!sym(dmg_column) > threshold) %>% 
    summarise(count = n_distinct(!!sym(fips_column)))
}

# Create a sequence of percentiles from 80th to 100th (in steps of 1%)
percentiles <- seq(0.80, 1.00, by = 0.01)

# Apply the function to each percentile and collect the results
treated_counts <- map_df(percentiles, count_treated_counties, df = flood_nonzero, dmg_column = "dmg_perEMP", fips_column = "FIPS_5")

# Add percentile information to the results
treated_counts <- treated_counts %>% 
  mutate(Percentile = percentiles * 100)

```

```{r eval=FALSE}

# Define a function to count unique counties and get damage per employee at a percentile threshold
count_treated_counties_with_damage <- function(percentile, df, dmg_column, fips_column) {
  threshold <- quantile(df[[dmg_column]], probs = percentile, na.rm = TRUE)
  count <- df %>% 
    filter(!!sym(dmg_column) > threshold) %>% 
    summarise(count = n_distinct(!!sym(fips_column))) %>% 
    pull(count)
  
  data.frame(
    Percentile = percentile * 100,
    DamagePerEmployee = threshold,
    Count = count
  )
}

# Create a sequence of percentiles from 80th to 100th (in steps of 1%)
percentiles <- seq(0.80, 1.00, by = 0.01)

# Apply the function to each percentile and collect the results
treated_counts <- map_df(percentiles, count_treated_counties_with_damage, df = flood_nonzero, dmg_column = "dmg_perEMP", fips_column = "FIPS_5")

# Print the resulting table
print(treated_counts)


```

# Treatment Variable

Based on the analysis above, the 95th percentile of flooding damage per employee (after dropping zeroes) is: 600.93156

ChatGPT again...

## Flood Threshold

```{r}

# Define the threshold for a serious flood (replace 'your_threshold' with an actual value)
serious_flood_threshold <- 600.93156

# Prepare the data
flood_df <- flood_df %>%
  # Define a serious flood
  mutate(serious_flood = if_else(dmg_perEMP >= serious_flood_threshold, 1, 0)) %>%
  # Group by county
  group_by(FIPS_5) %>%
  # Create the treatment variable
  mutate(treatment = cummax(serious_flood)) %>%
  ungroup()

```

## group variable

```{r}

flood_df <- flood_df %>%
  arrange(FIPS_5, year) %>%
  group_by(FIPS_5) %>%
  mutate(
    first_flood_year = ifelse(any(serious_flood == 1), min(year[serious_flood == 1]), NA_real_),
    group = ifelse(is.na(first_flood_year), 0, first_flood_year)
  ) %>%
  ungroup()




```

## Check

### Miss Counts

```{r eval=FALSE}

test_df %>% group_by(FIPS_5) %>% summarise(treated = max(serious_flood)) %>% filter(treated == 1) %>% nrow()

```

```{r eval=FALSE}

treat_count <- tibble(
  treated = test_df %>% group_by(FIPS_5) %>% summarise(treated = max(serious_flood)) %>% filter(treated == 1) %>% nrow(),
  untreated = test_df %>% group_by(FIPS_5) %>% summarise(treated = max(serious_flood)) %>% filter(treated == 0) %>% nrow()
)

```

### Group Variable

```{r}

check_group_assignments <- function(df, group_column, fips_column) {
  # Calculate the count of counties for each unique value in the group column
  group_counts <- df %>%
    group_by(!!sym(group_column)) %>%
    summarise(Count = n_distinct(!!sym(fips_column))) %>%
    ungroup()

  # Calculate the number of missing (NA) values in the group column
  missing_count <- sum(is.na(df[[group_column]]))

  # Calculate the number of counties with zero in the group column
  zero_count <- sum(df[[group_column]] == 0, na.rm = TRUE)

  # Combine the results into a single object
  list(
    group_counts = group_counts,
    missing_count = missing_count,
    zero_count = zero_count
  )
}

```

```{r}

# Assuming your dataframe is named flood_df and has the columns 'group' and 'FIPS_5'
check_group_assignments(flood_df, group_column = "group", fips_column = "FIPS_5")

```

# Missing Data

Try to get ChatGPT to help you with this! Either analyzing the data itself, or giving you code. The tricky thing is, you need to look at the data as a panel - so, which *counties* are missing a lot of data.

Previously (and you should be able to find code for this), when ChatGPT was working on the factor model, it did some analysis and imputation to deal with missing data, so maybe start there!

## % Missing

ChatGPT code!

```{r eval=FALSE}

# Creating a column for the percentage of missing data per county
miss_flood <- flood_df %>%
  mutate(missing_values = rowSums(is.na(.))) %>%
  mutate(missing_percentage = missing_values / ncol(.) * 100) %>%
  group_by(FIPS_5) %>%
  mutate(cty_miss = mean(missing_percentage)) %>%
  ungroup()

```

```{r eval=FALSE}

# Creating a column for the percentage of missing data per county, excluding Firm Death variables
excluded_variables <- c('firmdeath_firms', 'firmdeath_estabs', 'firmdeath_emp')
miss_flood_x <- miss_flood %>%
  select(-all_of(excluded_variables)) %>%
  mutate(missing_values_excluded = rowSums(is.na(.))) %>%
  mutate(missing_percentage_excluded = missing_values_excluded / ncol(.) * 100) %>%
  group_by(FIPS_5) %>%
  mutate(cty_miss_noFirm = mean(missing_percentage_excluded)) %>%
  ungroup()

# Adding the excluded data column back to the original dataframe
miss_flood$cty_miss_noFirm <- miss_flood_x$cty_miss_noFirm


```

# Analysis

## Variables

Variables needed for analysis

```{r}

# FIPS as Numeric (for 'did' package)
flood_df <- flood_df %>% mutate(
  FIPS = as.numeric(FIPS_5),
  ln_emp = log(emp + 1)
  )

```

## Filter missing

```{r}

# Function to drop counties with any missing values in a specified variable
drop_miss_cty <- function(data, variable) {
  # Identifying counties with any missing values in the specified variable
  counties_with_missing <- data %>%
    group_by(FIPS_5) %>%
    filter(any(is.na(.data[[variable]]))) %>%
    summarise() %>%
    pull(FIPS_5)

  # Excluding these counties from the dataset
  data_filtered <- data %>%
    filter(!(FIPS_5 %in% counties_with_missing))

  return(data_filtered)
}

# Example usage: 
# flood_BDS_filtered <- drop_counties_with_missing(flood_BDS, 'emp')


```

```{r}

# Function to verify the results of drop_miss_cty
verify_filtering <- function(original_data, filtered_data, variable) {
  # Check for missing values in the specified variable in the filtered dataset
  missing_values_check <- sum(is.na(filtered_data[[variable]]))

  # Compare the number of unique counties before and after filtering
  num_counties_before <- length(unique(original_data$FIPS_5))
  num_counties_after <- length(unique(filtered_data$FIPS_5))

  # Output the results
  cat("Verification Results:\n")
  cat("Missing values in '", variable, "' after filtering: ", missing_values_check, "\n", sep = "")
  cat("Number of counties before filtering: ", num_counties_before, "\n")
  cat("Number of counties after filtering: ", num_counties_after, "\n")
  cat("Number of counties removed: ", (num_counties_before - num_counties_after), "\n")

  # List of removed counties
  if (num_counties_before > num_counties_after) {
    removed_counties <- setdiff(unique(original_data$FIPS_5), unique(filtered_data$FIPS_5))
    cat("Counties removed:\n")
    print(removed_counties)
  }
}

# Example usage:
# verify_filtering(flood_BDS, flood_BDS_filtered, 'emp')


```

### emp

```{r}

emp_flood_df <- drop_miss_cty(flood_df, 'emp')

```

Check the cleaning:

```{r}

verify_filtering(flood_df, emp_flood_df, 'emp')

```

### estabs_exit_rate

```{r eval=FALSE}

estXr_flood_df <- drop_miss_cty(flood_df, 'estabs_exit_rate')

```

Check the cleaning:

```{r eval=FALSE}

verify_filtering(flood_df, estXr_flood_df, 'estabs_exit_rate')

```

# TWFE (ChatGPT)

I wanted to capture ChatGPT's attempt at a TWFE, even if I don't use it

```{r eval=FALSE}

library(plm)

# Preparing the data for the regression
flood_BDS_p <- pdata.frame(flood_BDS_sorted, index = c("FIPS_5", "year"))

# Creating the treatment variable: 1 if post-treatment period and 0 otherwise
flood_BDS_p$treatment <- as.integer(flood_BDS_p$year >= flood_BDS_p$group & flood_BDS_p$group > 0)

# Running the two-way fixed effects regression
# Employment as dependent variable, treatment as independent variable, controlling for county and year fixed effects
model <- plm(emp ~ treatment, data = flood_BDS_p, model = "within", effect = "twoways")

# Displaying the results
summary(model)


```

# Frontiers of DiD

From Causal Mixtape - Frontiers of DiD

```{r eval=FALSE}

library(did)

```

## Emp Model 1

```{r eval=FALSE}

emp_1 <- att_gt(yname="emp",
                  tname="year",
                  idname="FIPS",
                  gname="group",
                  data=emp_flood_df,
                  control_group = "notyettreated")

```

```{r eval=FALSE}

  emp_1_es <- aggte(emp_1, type="dynamic")

```

```{r eval=FALSE}

  ggdid(emp_1_es, xgap = 5)

```

```{r eval=FALSE}

  emp_1_es10 <- aggte(emp_1, type="dynamic", min_e = -10, max_e = 10)

```

```{r eval=FALSE}

  ggdid(emp_1_es10, xgap = 2)

```

```{r eval=FALSE}

  emp_1_overall <- aggte(emp_1, type="group")

```

```{r eval=FALSE}

  summary(emp_1_overall)

```

## Log Emp Model 1

```{r eval=FALSE}

ln_emp_1 <- att_gt(yname="ln_emp",
                  tname="year",
                  idname="FIPS",
                  gname="group",
                  data=emp_flood_df,
                  control_group = "notyettreated")

```

```{r eval=FALSE}

  ln_emp_1_es <- aggte(ln_emp_1, type="dynamic")

```

```{r eval=FALSE}

  ggdid(ln_emp_1_es, xgap = 4)

```

```{r eval=FALSE}

  ln_emp_1_es10 <- aggte(ln_emp_1, type="dynamic", min_e = -10, max_e = 10)

```

```{r eval=FALSE}

  ggdid(ln_emp_1_es10, xgap = 2)

```

```{r eval=FALSE}

  ln_emp_1_overall <- aggte(ln_emp_1, type="group")
  summary(ln_emp_1_overall)

```

## estabs_exit_rate

```{r eval=FALSE}

est_ex_1 <- att_gt(yname="estabs_exit_rate",
                  tname="year",
                  idname="FIPS",
                  gname="group",
                  data=estXr_flood_df,
                  control_group = "notyettreated")

```

```{r eval=FALSE}

  est_ex_1_es <- aggte(est_ex_1, type="dynamic")

```

```{r eval=FALSE}

  ggdid(est_ex_1_es, xgap = 5)

```

```{r eval=FALSE}

  est_ex_1_overall <- aggte(est_ex_1, type="group")

```

```{r eval=FALSE}

  summary(est_ex_1_overall)

```

# Data Exploration

```{r eval=FALSE}

flood_df %>% group_by(st) %>% count(st)

```