---
title: "SHELDUS Deryugina Analysis"
author: "Sam Neylon"
date: '2024-08-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(stringr)
library(did)

# library(scales)

set.seed(7008)

```



# Notes

##(8-17-24)

### My goals:
  1. Dery Outcomes + Dery Treats + SN Methods: See if I get the same results as Deryugina, using her outcomes, her hurricanes (wind speed), but the methods I am using.
    1.a: Maybe don't use covariates at first, which will give you different outcomes than Deryugina potentially.
  2. Dery Outcomes + SHELDUS Treats + SN Methods: See how the outcomes change using the SHELDUS hurricanes (any hurricane damage).
  3. Dery Covariates + Dery Outcomes + Dery Treats + SN Methods: Add in the covariates!
  
Potential:
  * BDS outcomes?: Should I experiment with seeing how the BDS results differ from Deryugina
  * Wind speed and Damage: Could be interesting!
  
Other Goals:

  * Figure out how to estimate effects for specific time periods - 10 years, 5 years - inside the 'did' and other packages.
  
  
##(8-18-24)

• 1979 - many more SHELDUS hurricanes
		○ 526 counties affected
			§ Next greatest is 85!
		!!!!! I created a version of the data without 1979! If I want to change it back, I need to change it everywhere below!

• BIG NOTE: 1979 wasn't really the issue…
			§ Or at least, another year became a big year (but ~350 counties, not 500 counties)
			§ Issue goes deeper

• Fixes for SHELDUS estimates - should I reverse these?
		○ Changed 'est_method' to "reg"
			§ This was because I think the IPW part of the doubly-robust estimation was the glm that wasn't converging.
		○ Na.rm for aggte()
			§ For the SHELDUS estimation, I added 'na.rm', but I'd rather figure out why there is missing?

• Interesting finding:
		○ There is an employment effect for damage, but not for wind speed!

## Deryugina Analysis Notes:

NOTE: Deryugina uses spatial OLS for her regressions.

- She gets rid of wind_speed in her cleaned data.

# Import 79

## 1979!

The SHELDUS data in 1979 is odd - splitting damage between many more counties (526) than any other year (the next biggest year is 85 counties).

Until I can figure out what the deal is, I am going create a special version of the data which calculates the first hurricane year without 1979.

I have created this different dataset: SHELDUS_Deryugina79.rds

```{r}

df <- readRDS(here("data/SHELDUS/SHELDUS_Deryugina79.rds"))

```



# Dery Outcomes + Dery Treats + SN Methods

Dery Outcomes + Dery Treats + SN Methods: See if I get the same results as Deryugina, using her outcomes, her hurricanes (wind speed), but the methods I am using.

## Govt Transfers

```{r eval=FALSE}

gov_t_D <- att_gt(yname="log_curr_trans_ind_gov_pc",
                  tname="year",
                  idname="county_fips",
                  gname="group_Dery",
                  data=df,
                  control_group = "notyettreated",
                  cores = 4)

```

```{r eval=FALSE}

  gov_t_D_es10 <- aggte(gov_t_D, type="dynamic", min_e = -10, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(gov_t_D_es10)

```

```{r eval=FALSE}

  ggdid(gov_t_D_es10, xgap = 2)

```

```{r eval=FALSE}

  gov_t_D_att10 <- aggte(gov_t_D, type="dynamic", min_e = 1, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(gov_t_D_att10)

```

## Biz Transfers

NOTE: 8045 rows dropped for missing data. This mostly aligns with Deryugina descriptives (pg 37), even though there are less missing than there should be (I probably dropped some during cleaning):

Per capita transfers from gov. obvs: 49,245
Per capita transfers from bus. obvs: 40,027

```{r eval=FALSE}

biz_t_D <- att_gt(yname="log_curr_trans_ind_bus_pc",
                  tname="year",
                  idname="county_fips",
                  gname="group_Dery",
                  data=df,
                  control_group = "notyettreated",
                  cores = 4)

```

```{r eval=FALSE}

  biz_t_D_es10 <- aggte(biz_t_D, type="dynamic", min_e = -10, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(biz_t_D_es10)

```

```{r eval=FALSE}

  ggdid(biz_t_D_es10, xgap = 2)

```

```{r eval=FALSE}

  biz_t_D_att10 <- aggte(biz_t_D, type="dynamic", min_e = 1, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(biz_t_D_att10)

```

## Emp Rate (Total Adult Population)

```{r eval=FALSE}

emp_tot_adult_D <- att_gt(yname="emp_rate_tot_adult",
                  tname="year",
                  idname="county_fips",
                  gname="group_Dery",
                  data=df,
                  control_group = "notyettreated",
                  cores = 4)

```

```{r eval=FALSE}

  emp_tot_adult_D_es10 <- aggte(emp_tot_adult_D, type="dynamic", min_e = -10, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(emp_tot_adult_D_es10)

```

```{r eval=FALSE}

  ggdid(emp_tot_adult_D_es10, xgap = 2)

```

```{r eval=FALSE}

  emp_tot_adult_D_att10 <- aggte(emp_tot_adult_D, type="dynamic", min_e = 1, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(emp_tot_adult_D_att10)

```

## Avg Wages

```{r eval=FALSE}

avg_wage_D <- att_gt(yname="log_avg_wage_sal_disb",
                  tname="year",
                  idname="county_fips",
                  gname="group_Dery",
                  data=df,
                  control_group = "notyettreated",
                  cores = 4)

```

```{r eval=FALSE}

  avg_wage_D_es10 <- aggte(avg_wage_D, type="dynamic", min_e = -10, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(avg_wage_D_es10)

```

```{r eval=FALSE}

  ggdid(avg_wage_D_es10, xgap = 2)

```

```{r eval=FALSE}

  avg_wage_D_att10 <- aggte(avg_wage_D, type="dynamic", min_e = 1, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(avg_wage_D_att10)

```

# SHELDUS Hurricanes + Dery Outcomes + Dery Treats + SN Methods

SHELDUS Hurricanes + Dery Outcomes + Dery Treats + SN Methods: Using Deryugina outcomes, SHELDUS hurricanes (any hurricane damage), but the methods I am using.

## Govt Transfers

```{r eval=FALSE}

gov_t_SH <- att_gt(yname="log_curr_trans_ind_gov_pc",
                  tname="year",
                  idname="county_fips",
                  gname="group_SHLD",
                  data=df,
                  control_group = "notyettreated",
                  cores = 4,
                  pl = TRUE,
                  est_method = "reg")

```

```{r eval=FALSE}

  gov_t_SH_es10 <- aggte(gov_t_SH, type="dynamic", min_e = -10, max_e = 10, balance_e = 10, na.rm = TRUE)

```

```{r eval=FALSE}

summary(gov_t_SH_es10)

```

```{r eval=FALSE}

  ggdid(gov_t_SH_es10, xgap = 2)

```

```{r eval=FALSE}

  gov_t_SH_att10 <- aggte(gov_t_SH, type="dynamic", min_e = 1, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(gov_t_SH_att10)

```

## Biz Transfers

```{r eval=FALSE}

biz_t_SH <- att_gt(yname="log_curr_trans_ind_bus_pc",
                  tname="year",
                  idname="county_fips",
                  gname="group_SHLD",
                  data=df,
                  control_group = "notyettreated",
                  cores = 4)

```

```{r eval=FALSE}

  biz_t_SH_es10 <- aggte(biz_t_SH, type="dynamic", min_e = -10, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(biz_t_SH_es10)

```

```{r eval=FALSE}

  ggdid(biz_t_SH_es10, xgap = 2)

```

```{r eval=FALSE}

  biz_t_SH_att10 <- aggte(biz_t_SH, type="dynamic", min_e = 1, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(biz_t_SH_att10)

```

## Emp Rate (Total Adult Population)

```{r eval=FALSE}

emp_tot_adult_SH <- att_gt(yname="emp_rate_tot_adult",
                  tname="year",
                  idname="county_fips",
                  gname="group_SHLD",
                  data=df,
                  control_group = "notyettreated",
                  cores = 4,
                  pl = TRUE,
                  est_method = "reg")

```

```{r eval=FALSE}

  emp_tot_adult_SH_es10 <- aggte(emp_tot_adult_SH, type="dynamic", min_e = -10, max_e = 10, balance_e = 10, na.rm = TRUE)

```

```{r eval=FALSE}

summary(emp_tot_adult_SH_es10)

```

```{r eval=FALSE}

  ggdid(emp_tot_adult_SH_es10, xgap = 2)

```

```{r eval=FALSE}

  emp_tot_adult_SH_att10 <- aggte(emp_tot_adult_SH, type="dynamic", min_e = 1, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(emp_tot_adult_SH_att10)

```

## Avg Wages

```{r eval=FALSE}

avg_wage_SH <- att_gt(yname="log_avg_wage_sal_SHisb",
                  tname="year",
                  idname="county_fips",
                  gname="group_SHLD",
                  data=df,
                  control_group = "notyettreated",
                  cores = 4)

```

```{r eval=FALSE}

  avg_wage_SH_es10 <- aggte(avg_wage_SH, type="dynamic", min_e = -10, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(avg_wage_SH_es10)

```

```{r eval=FALSE}

  ggdid(avg_wage_SH_es10, xgap = 2)

```

```{r eval=FALSE}

  avg_wage_SH_att10 <- aggte(avg_wage_SH, type="dynamic", min_e = 1, max_e = 10, balance_e = 10)

```

```{r eval=FALSE}

summary(avg_wage_SH_att10)

```











# Dery Covariates + Dery Outcomes + Dery Treats + SN Methods

## Deryugina Equation 1

[Not the version she uses in the paper, which is a spatial regression]

* Note: if you do not have ols_spatial_HAC, you can obtain the paper's point estimates with standard errors clustered by county by running the following code (currently commented out)

* reghdfe `outcome' hurricane hurr_*, absorb(i.county_fips i.year#i.coastal i.year#c.land_area1970 i.year#c.log_pop1969 i.year#c.frac_young1969 i.year#c.frac_old1969 i.year#c.frac_black1969 i.year#c.log_wage_pc1969 i.year#c.emp_rate1969) vce(cluster county_fips)

# Data Exploration

## Descriptives

### Continuous

```{r eval=FALSE}

library(scales)

```

```{r eval=FALSE}

cont_desc_fn <- function(data, variables) {
  stats <- data %>%
    summarise(across(all_of(variables), 
                    list(N = ~sum(!is.na(.)),
                          Mean = ~round(mean(., na.rm = TRUE), 2),
                          P25 = ~round(quantile(., 0.25, na.rm = TRUE), 2),
                          P50 = ~round(quantile(., 0.50, na.rm = TRUE), 2),
                          P75 = ~round(quantile(., 0.75, na.rm = TRUE), 2),
                          P95 = ~round(quantile(., 0.95, na.rm = TRUE), 2),
                          Max = ~round(max(., na.rm = TRUE), 2)),
                     .names = "{.col}__{.fn}"))
  
  stats_long <- stats %>%
    pivot_longer(cols = everything(), names_to = c("Variable", "Statistic"), 
                 names_sep = "__") %>%
    pivot_wider(names_from = Statistic, values_from = value)
  
  return(stats_long)
}

# Example usage
# Assuming df is your data frame and you have a list of variables you want to summarize
#cont_vars <- c("emp", "ln_emp")
cont_vars_noZero <- c("PropertyDmg_Adj2022", "PropertyDmgPerCapita_Adj2022")

cont_tbl_nZ <- df %>% 
  filter(PropertyDmg_Adj2022 > 0) %>% 
  cont_desc_fn(data = ., variables = cont_vars_noZero)

#cont_tbl <- cont_desc_fn(ln_emp_flood_df, cont_vars)

#cont_tbl_full <- bind_rows(cont_tbl, cont_tbl_nZ)
cont_tbl_full <- cont_tbl_nZ

cont_tbl_full <- cont_tbl_full %>%
  mutate(across(-c(Variable, N), ~case_when(
    #Variable == "emp" ~ paste0(comma(. / 1000, accuracy = 0.01), "k"),
    #Variable == "ln_emp" ~ paste0(comma(., accuracy = 0.01)),
    Variable == "PropertyDmg_Adj2022" ~ paste0("$", comma(. / 1000, accuracy = 0.01), "k"),
    Variable == "PropertyDmgPerCapita_Adj2022" ~ paste0("$", comma(., accuracy = 0.01)))
  ))

```

```{r eval=FALSE}

# Print the summary table
print(cont_tbl_full)

```

###1979

```{r eval=FALSE}

cont_desc_fn <- function(data, variables) {
  stats <- data %>%
    summarise(across(all_of(variables), 
                    list(N = ~sum(!is.na(.)),
                          Mean = ~round(mean(., na.rm = TRUE), 2),
                          P25 = ~round(quantile(., 0.25, na.rm = TRUE), 2),
                          P50 = ~round(quantile(., 0.50, na.rm = TRUE), 2),
                          P75 = ~round(quantile(., 0.75, na.rm = TRUE), 2),
                          P95 = ~round(quantile(., 0.95, na.rm = TRUE), 2),
                          Max = ~round(max(., na.rm = TRUE), 2)),
                     .names = "{.col}__{.fn}"))
  
  stats_long <- stats %>%
    pivot_longer(cols = everything(), names_to = c("Variable", "Statistic"), 
                 names_sep = "__") %>%
    pivot_wider(names_from = Statistic, values_from = value)
  
  return(stats_long)
}

# Example usage
# Assuming df is your data frame and you have a list of variables you want to summarize
#cont_vars <- c("emp", "ln_emp")
cont_vars_noZero <- c("PropertyDmg_Adj2022", "PropertyDmgPerCapita_Adj2022")

cont_tbl_nZ <- df %>% 
  filter(group_SHLD == 1979 & year == 1979) %>% 
  filter(PropertyDmg_Adj2022 > 0) %>% 
  cont_desc_fn(data = ., variables = cont_vars_noZero)

#cont_tbl <- cont_desc_fn(ln_emp_flood_df, cont_vars)

#cont_tbl_full <- bind_rows(cont_tbl, cont_tbl_nZ)
cont_tbl_full <- cont_tbl_nZ

cont_tbl_full <- cont_tbl_full %>%
  mutate(across(-c(Variable, N), ~case_when(
    #Variable == "emp" ~ paste0(comma(. / 1000, accuracy = 0.01), "k"),
    #Variable == "ln_emp" ~ paste0(comma(., accuracy = 0.01)),
    Variable == "PropertyDmg_Adj2022" ~ paste0("$", comma(. / 1000, accuracy = 0.01), "k"),
    Variable == "PropertyDmgPerCapita_Adj2022" ~ paste0("$", comma(., accuracy = 0.01)))
  ))

```

```{r eval=FALSE}

# Print the summary table
print(cont_tbl_full)

```

## 1979 SHELDUS Issue

```{r eval=FALSE}

# 1979 exploration

SH_79 <- df %>%
  #filter(group_SHLD %in% c(1979, 1980) & year %in% c(1979, 1980)) %>% 
  filter(year %in% c(1979, 1980, 1981)) %>% 
  select(FIPS_5, state_name, county_name_SHELDUS, PropertyDmg_Adj2022, PropertyDmgPerCapita_Adj2022, county_fips, year, state_fips, hurricane, central_wind_speed)

write_csv(SH_79, here("data/data_check/SHELD-Dery_1979.csv"))

```

## County Count Check

How many counties per year?

```{r eval=FALSE}

# Assuming your data frame is named 'df'
df %>%
  # Group by the 'year' variable
  group_by(year) %>%
  # Count the number of distinct counties (FIPS_5) in each year
  summarize(county_count = n_distinct(FIPS_5))

```

## Group Counts

### Deryugina

```{r eval=FALSE}

df %>%
  # Group by the 'group_SHLD' variable
  group_by(group_Dery) %>%
  # Count the number of counties in each 'group_SHLD' year
  summarize(county_count = n_distinct(FIPS_5))

```

### SHELDUS

```{r eval=FALSE}

df %>%
  # Group by the 'group_SHLD' variable
  group_by(group_SHLD) %>%
  # Count the number of counties in each 'group_SHLD' year
  summarize(county_count = n_distinct(FIPS_5))

```

## Non-Convergence

Get a data set saved that you can put into ChatGPT, to figure out why your data might not be converging.

```{r eval=FALSE}

non_converge_df <- df %>% select(year, county_fips, state_fips, group_SHLD, any_hurrDMG_SHLD, PropertyDmgPerCapita_Adj2022, log_curr_trans_ind_gov_pc)

```

```{r eval=FALSE}

write_csv(non_converge_df, here("data/data_check/non_converge.csv"))

```