---
title: "Forecast Zones"
author: "Sam Neylon"
date: '2023-11-05'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(stringr)

```

# Notes
##(11-5-23)

Okay, so I figured out one way to do this. I do a join which is way too big, with every combination (hopefully not permutation...) of Forecast Zone and county.
  * Then I group by county, and do a 0 or 1 depending on if a hurricane hit.

Episode ID: One interesting wrinkle, because hurricanes have an episode ID, I can group by the HURRICANE itself... could be an interesting way to do the treatment effects...
  * Also, in case multiple hurricanes hit in a year, I can see this.

!!!NOTE!!! -- You should go through the code below, and clean up any joins that do weird "FIPS.y" type stuff - you want to make sure your joins are predictable (and there isn't duplication of column names!)

# Import Zones

Zones are in a .dbx file, which is apparently an Outlook database? But when I look at the file, they are delimited with a vertical bar "|" - so I can use readr to read it in (maybe).

File downloaded from here: https://www.weather.gov/gis/ZoneCounty

```{r}

zone_county <- read_delim(here("data/Labels/forecastZones_counties.dbx"), delim = "|", col_names = c(
  "STATE",
  "ZONE",
  "CWA",
  "NAME",
  "STATE_ZONE",
  "COUNTY",
  "FIPS",
  "TIME_ZONE",
  "FE_AREA",
  "LAT",
  "LON"
))



```

## Combo ID

Combine FIPS and Zone numbers

```{r}

zone_county <- zone_county %>% mutate(
  fips_zone = str_c(FIPS, ZONE)
)

```


# NWS Events Data

```{r}

nws_detail_22 <- read_csv(here("data/storm_events/test/StormEvents_details-ftp_v1.0_d2022_c20231017.csv"))

```

## State FIPS

```{r}

state_fips <- read_csv(here("data/Labels/state_fips.csv")) %>% 
  mutate(fips_num = as.numeric(FIPS))

```

```{r}

nws_detail_22 <- left_join(nws_detail_22, state_fips, join_by(STATE_FIPS == fips_num))

```

## Full FIPS

```{r}

nws_detail_22 <- nws_detail_22 %>% mutate(
  full_FIPS = case_when(
    str_length(CZ_FIPS) == 3  ~ as.character(CZ_FIPS),
    str_length(CZ_FIPS) == 2  ~ str_c("0", as.character(CZ_FIPS)),
    str_length(CZ_FIPS) == 1  ~ str_c("00", as.character(CZ_FIPS))
  ),
  STATE_ZONE = str_c(Postal, full_FIPS)
)

```


```{r}

zoneJoin <- inner_join(zone_county, filter(nws_detail_22, CZ_TYPE == "Z"), join_by(STATE_ZONE == STATE_ZONE))

```

## Data Check

```{r eval=FALSE}

tbl_zoneJoin <- zoneJoin %>% group_by(STATE_ZONE) %>% count() %>% mutate(STATE = str_sub(STATE_ZONE, 1, 2))

```

```{r eval=FALSE}

tbl2_zoneJoin <- zoneJoin %>% group_by(fips_zone) %>% count()

```

```{r eval=FALSE}

ev_tbl <- zoneJoin %>% filter(CZ_TYPE == "Z") %>% count(EVENT_ID)

```


# Hurricane by County

```{r}

county_hurr <- zoneJoin %>% mutate(
  hurr = ifelse(EVENT_TYPE == "Hurricane", 1, 0)
) %>% 
  group_by(FIPS.y) %>% 
  summarise(
    hurr_sum = sum(hurr)
  ) %>% 
  mutate(hurr = ifelse(hurr_sum > 0, 1, 0))

```


## Data Check

```{r eval=FALSE}

hurr2_nws <- zoneJoin %>% filter(EVENT_TYPE == "Hurricane")

```



# Tableau Version

```{r eval=FALSE}

write_csv(zoneJoin, here("data/storm_events/test/zoneJoin_2022.csv"))

```

```{r eval=FALSE}

write_csv(county_hurr, here("data/storm_events/test/county_hurr_2022.csv"))

```


# Zone Data Check

Make list of zones that lie in more than one county.

```{r eval=FALSE}

tbl_zone <- zone_county %>% group_by(STATE_ZONE) %>% count() %>% mutate(STATE = str_sub(STATE_ZONE, 1, 2))

```

```{r eval=FALSE}

write_csv(tbl_zone, here("data/Labels/working_data/forecast_zones_count.csv"))

```

```{r eval=FALSE}

st_tbl_zone <- tbl_zone %>% group_by(STATE) %>% 
  summarise(single = sum(ifelse(n == 1, 1, 0)),
            multi = sum(ifelse(n > 1, 1, 0)))

```

