---
title: "NWS Flooding Clean"
output: html_document
date: "2023-12-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(stringr)

```

# Notes
##(12-28-23)

NOTE: There is hinky stuff with the geo join (columns with the same names, etc), so clean that up if you end up using that file for anything but Tableau.

# NWS Events Data

```{r}

nws_detail_22 <- read_csv(here("data/storm_events/test/StormEvents_details-ftp_v1.0_d2022_c20231017.csv"))

```

## State FIPS

```{r}

state_fips <- read_csv(here("data/Labels/state_fips.csv")) %>% 
  mutate(fips_num = as.numeric(FIPS))

```

```{r}

nws_detail_22 <- left_join(nws_detail_22, state_fips, join_by(STATE_FIPS == fips_num))

```

## County Only

```{r}

nws_counties <- nws_detail_22 %>% filter(CZ_TYPE == "C")

```


## Full FIPS

```{r}

nws_counties <- nws_counties %>% mutate(
  full_FIPS = case_when(
    str_length(CZ_FIPS) == 3  ~ as.character(CZ_FIPS),
    str_length(CZ_FIPS) == 2  ~ str_c("0", as.character(CZ_FIPS)),
    str_length(CZ_FIPS) == 1  ~ str_c("00", as.character(CZ_FIPS))
  ),
  FIPS_5 = str_c(FIPS, full_FIPS)
)

```

## Just Floods

```{r}

nws_flood <- nws_counties %>% filter(EVENT_TYPE == "Flood" | EVENT_TYPE == "Flash Flood")

```

## Convert Damage to Number

```{r}

nws_flood <- nws_flood %>% 
  mutate(
    PROP_DMG = case_when(
      is.na(DAMAGE_PROPERTY) ~ NA_real_,
      str_detect(DAMAGE_PROPERTY, "K") ~ parse_number(DAMAGE_PROPERTY) * 1000,
      str_detect(DAMAGE_PROPERTY, "M") ~ parse_number(DAMAGE_PROPERTY) * 1000000,
      TRUE ~ NA_real_
    )
  ) %>% 
  relocate(PROP_DMG, .before = DAMAGE_PROPERTY)
  

```

# Drop Raw Data

```{r eval=FALSE}

rm(nws_detail_22, nws_counties)

```



# Tableau Version

```{r eval=FALSE}

write_csv(nws_counties, here(""))

```


# Data Exploration

```{r section-setup}
knitr::opts_chunk$set(eval = FALSE)

```

Note: This will not stop the chunks from running when I use "Run All", but I thought I would keep it here, in case I end up knitting the document for some reason?

## Explore

```{r eval=FALSE}

nws_flood %>% group_by(EVENT_TYPE) %>% count()

```

```{r eval=FALSE}

nws_detail_22 %>% group_by(EVENT_TYPE) %>% count()

```


```{r eval=FALSE}



```


## End Section

```{r end-section}
knitr::opts_chunk$set(eval = TRUE)

```

# Geo Data 

Add County names to data, so I can get them in Tableau

```{r}

fips_df <- read_csv(here("data/Labels/FIPS_all_counties.csv"))

```


```{r}

geo_nws_flood <- left_join(nws_flood, fips_df, join_by(FIPS_5 == FIPS_text))

```


## Tableau Geo Version

```{r eval=FALSE}

write_csv(geo_nws_flood, here("data/storm_events/test/flood_22_geo.csv"))

```



# County Sums

```{r eval=FALSE}

county_flood <- nws_flood %>% group_by(FIPS_5) %>% 
  summarise(prop_dmg = sum(PROP_DMG, na.rm = TRUE))

```

