---
title: "SHELDUS Import"
output: html_document
date: "2024-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(stringr)

set.seed(1157)

```

# Notes

##(3-28-24)

Importing SHELDUS data. A lot of missing data and other analysis still to go through!

##(4-2-24)

I used ChatGPT to analyze just the hurricane damage data, and it seemed clean and nothing missing!




# Import SHELDUS

```{r}

SHELDUS_df <- read_csv(here("data/SHELDUS/sheldus_1985-2022_hurr-flood-wind.csv"),
                       col_types = cols(
  `State Name` = col_character(),
  `County Name` = col_character(),
  `County FIPS` = col_character(),
  Hazard = col_character(),
  Year = col_double(),
  Month = col_double(),
  CropDmg = col_double(),
  `CropDmg(ADJ 2022)` = col_double(),
  `CropDmgPerCapita(ADJ 2022)` = col_double(),
  PropertyDmg = col_double(),
  `PropertyDmg(ADJ 2022)` = col_double(),
  `PropertyDmgPerCapita(ADJ 2022)` = col_double(),
  Injuries = col_double(),
  InjuriesPerCapita = col_double(),
  Fatalities = col_double(),
  FatalitiesPerCapita = col_double(),
  Duration_Days = col_double(),
  Fatalities_Duration = col_double(),
  Injuries_Duration = col_double(),
  Property_Damage_Duration = col_double(),
  Crop_Damage_Duration = col_double(),
  Records = col_double())
) %>% 
  rename(
  state_name = `State Name`,
  county_name = `County Name`,
  county_FIPS = `County FIPS`,
  CropDmg_Adj2022 = `CropDmg(ADJ 2022)`,
  CropDmgPerCapita_Adj2022 = `CropDmgPerCapita(ADJ 2022)`,
  PropertyDmg_Adj2022 = `PropertyDmg(ADJ 2022)`,
  PropertyDmgPerCapita_Adj2022 = `PropertyDmgPerCapita(ADJ 2022)`
  ) %>% 
  mutate(
    county_FIPS = str_replace_all(county_FIPS, "'", "")
  )

```

## Hurricane Data

```{r}

hurr_SHELDUS <- SHELDUS_df %>% 
  select(state_name:Month | PropertyDmg:PropertyDmgPerCapita_Adj2022 | Duration_Days | Records) %>% 
  filter(Hazard == "Hurricane/Tropical Storm" & Year >= 1996)

```

## Sum damage by year

```{r}

hurr_SHELDUS <- hurr_SHELDUS %>%
  group_by(state_name, county_name, county_FIPS, Year) %>%
  summarise(
    PropertyDmg = sum(PropertyDmg, na.rm = TRUE),
    PropertyDmg_Adj2022 = sum(PropertyDmg_Adj2022, na.rm = TRUE),
    PropertyDmgPerCapita_Adj2022 = sum(PropertyDmgPerCapita_Adj2022, na.rm = TRUE),
    .groups = 'drop' # This option drops the grouping structure afterwards
  ) %>% 
  mutate(
    SHELDUS = 1
  )

```


## Export Hurricane Data

```{r eval=FALSE}

write_csv(hurr_SHELDUS, here("data/SHELDUS/hurr_SHELDUS_df.csv"))

```


# Join with BDS

```{r}

# Function to load the bds_df dataset
load_bds_df <- function(filepath) {
  read_csv(filepath, col_types = cols(
  year = col_double(),
  st = col_character(),
  cty = col_character(),
  firms = col_double(),
  estabs = col_double(),
  emp = col_double(),
  denom = col_double(),
  estabs_entry = col_double(),
  estabs_entry_rate = col_double(),
  estabs_exit = col_double(),
  estabs_exit_rate = col_double(),
  job_creation = col_double(),
  job_creation_births = col_double(),
  job_creation_continuers = col_double(),
  job_creation_rate_births = col_double(),
  job_creation_rate = col_double(),
  job_destruction = col_double(),
  job_destruction_deaths = col_double(),
  job_destruction_continuers = col_double(),
  job_destruction_rate_deaths = col_double(),
  job_destruction_rate = col_double(),
  net_job_creation = col_double(),
  net_job_creation_rate = col_double(),
  reallocation_rate = col_double(),
  firmdeath_firms = col_double(),
  firmdeath_estabs = col_double(),
  firmdeath_emp = col_double(),
  FIPS_5 = col_character()
  )) %>%
    filter(year >= 1986)
}

# Function to join datasets and filter based on State FIPS codes
join_and_filter <- function(bds_df, hurr_SHELDUS) {
  bds_df %>%
    left_join(hurr_SHELDUS, by = c("FIPS_5" = "county_FIPS", "year" = "Year")) %>%
    mutate(pre_96 = if_else(year < 1996, 1, 0),
           across(starts_with("PropertyDmg"), ~replace_na(.x, 0))
    ) %>% 
    select(!state_name:county_name)
}

```


## Run Functions

```{r}

# Load and process the datasets
bds_df_path <- here("data/BDS/bds_df.csv")

bds_df <- load_bds_df(bds_df_path)

```

## Join

```{r}

# Join the datasets and filter
hurrSHELDUS_bds_df <- join_and_filter(bds_df, hurr_SHELDUS)

```

### Export

```{r eval=FALSE}

write_csv(hurrSHELDUS_bds_df, here("data/SHELDUS/BDS_hurr_SHELDUS.csv"))

```

# Treatment Variable

Setting the hurricane damage threshold at 80th percentile for now (based on some analysis I did on ChatGPT).

```{r}

# Define serious hurricane damage at 80the percentile of SHELDUS counties

serious_threshold <- hurrSHELDUS_bds_df %>%
  filter(SHELDUS == 1) %>%  # Filter rows where SHELDUS == 1
  summarise(quantile_80 = quantile(PropertyDmgPerCapita_Adj2022, 0.8, na.rm = TRUE)) %>%
  pull(quantile_80)  # Extract the quantile value

# Check the calculated threshold
serious_threshold

```


```{r}

# Prepare the data
hurrSHELDUS_bds_df <- hurrSHELDUS_bds_df %>%
  # Define a serious flood
  mutate(serious_hurr = if_else(PropertyDmgPerCapita_Adj2022 >= serious_threshold, 1, 0)) %>%
  # Group by county
  group_by(FIPS_5) %>%
  # Create the treatment variable
  mutate(treatment = cummax(serious_hurr)) %>%
  ungroup()

```

## group variable

```{r}

hurrSHELDUS_bds_df <- hurrSHELDUS_bds_df %>%
  arrange(FIPS_5, year) %>%
  group_by(FIPS_5) %>%
  mutate(
    first_hurr_year = ifelse(any(serious_hurr == 1), min(year[serious_hurr == 1]), NA_real_),
    group = ifelse(is.na(first_hurr_year), 0, first_hurr_year)
  ) %>%
  ungroup()

```

# Other Data Management

## hurr_state variable

```{r}

hurr_states <-  read_csv(here("data/Labels/hurricane_states.csv"))

```

```{r}

hurrSHELDUS_bds_df <- hurrSHELDUS_bds_df %>% 
  left_join(hurr_states, by = c("st" = "FIPS"))

```

### Export Full Data

```{r}

# Full Data
write_csv(hurrSHELDUS_bds_df, here("data/SHELDUS/BDS_hurr_SHELDUS.csv"))

```

# Hurr States Filter

```{r}

# Just Hurricane States
hs_hurrSHELDUS_bds_df <- hurrSHELDUS_bds_df %>% 
  filter(hurr_state == 1)

```

### Export Hurr States Data

```{r}

# Hurricane States Data
write_csv(hs_hurrSHELDUS_bds_df, here("data/SHELDUS/hs_BDS_hurr_SHELDUS.csv"))

```

# Sample

```{r eval=FALSE}

sample_df <- hurrSHELDUS_bds_df %>%
  slice_sample(n = 100)

```

```{r eval=FALSE}

write_csv(sample_df, here("data/SHELDUS/sample_df.csv"))

```
