---
title: "SHELDUS Hurricane Import"
output: html_document
date: "2024-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
library(stringr)
library(haven)

set.seed(1157)

```

# Notes

##(3-28-24)

Importing SHELDUS data. A lot of missing data and other analysis still to go through!

##(4-2-24)

I used ChatGPT to analyze just the hurricane damage data, and it seemed clean and nothing missing!

##(4-3-24)

### Dropping 80th percentile for median
Okay, so, I have been doing analysis. I am not getting significant results with 80th percentile of damage, unless I drop Houston TX! While I don't want to data dredge, I do think it is legitimate to drop my serious threshold, to get more data. In fact, this should include more counties which didn't have an employment effect!

I am doing the median instead.

- Update: Median made it worse! So, I'm not data dredging haha. But Houston really is an outlier, I need to figure out how to deal with it.

##(4-29-24)

### Deryugina

Now this is a Deryugina script!

### All Hurricanes?

No more serious hurricanes nonsense - if there is damage from a hurricane, it's a hurricane!

##(6-9-24)

Because my earlier SHELDUS data didn't go back far enough, I got a new file from SHELDUS, going back to 1960, but aggregated by year. For the code dealing with 1985 onwards monthly data, see 'SHELDUS_Deryugina_85-22.Rmd'.

### To Do

DONE - Adjust SHELDUS treatment so it is only between 1979-2002
TO DO - Do Deryugina versions of the treatment variables

# Important Notes

## Treat Var Name Changes

I have added 'SHLD' and 'Dery' suffixes to treatment variables, to indicate whether they come from SHELDUS or Deryugina hurricane indicators.

I have changed the following variable names, which will need to be changed in the analysis code!

'treatment' -> 'post_treat'
'treat' -> 'ever_treated' 
'treatment_post' -> 'treat_DiD'

#(08-17-24)

##Different hurricane years!!

Some years where I calculate there was hurricane damage, there isn't this damage in Deryugina.

NEW NOTE! This is fine! I wanted to run the SHELDUS treatment models using the Deryugina outcomes!


# Import SHELDUS

```{r eval=TRUE}

SHELDUS_df <- read_csv(here("data/SHELDUS/sheldus_1960-2023_byYR_hurr-flood-wind.csv"),
                       col_types = cols(
  `State Name` = col_character(),
  `County Name` = col_character(),
  `County FIPS` = col_character(),
  Hazard = col_character(),
  Year = col_double(),
  CropDmg = col_double(),
  `CropDmg(ADJ 2022)` = col_double(),
  `CropDmgPerCapita(ADJ 2022)` = col_double(),
  PropertyDmg = col_double(),
  `PropertyDmg(ADJ 2022)` = col_double(),
  `PropertyDmgPerCapita(ADJ 2022)` = col_double(),
  Injuries = col_double(),
  InjuriesPerCapita = col_double(),
  Fatalities = col_double(),
  FatalitiesPerCapita = col_double(),
  Duration_Days = col_double(),
  Fatalities_Duration = col_double(),
  Injuries_Duration = col_double(),
  Property_Damage_Duration = col_double(),
  Crop_Damage_Duration = col_double(),
  Records = col_double()
)) %>% 
  rename(
  state_name = `State Name`,
  county_name_SHELDUS = `County Name`,
  FIPS_5 = `County FIPS`,
  PropertyDmg_Adj2022 = `PropertyDmg(ADJ 2022)`,
  PropertyDmgPerCapita_Adj2022 = `PropertyDmgPerCapita(ADJ 2022)`
  ) %>% 
  mutate(
    FIPS_5 = str_replace_all(FIPS_5, "'", "")
  )

```

## Hurricane Data

```{r}

hurr_SHELDUS <- SHELDUS_df %>% 
  select(state_name:Year | PropertyDmg:PropertyDmgPerCapita_Adj2022 | Duration_Days | Records) %>% 
  filter(Hazard == "Hurricane/Tropical Storm" & Year >= 1965 & Year <= 2012) %>% 
  mutate(
    SHELDUS = 1
  )

```

### Drop SHELDUS Data

```{r}

rm(SHELDUS_df)

```

## Export Hurricane Data

```{r eval=FALSE}

write_csv(hurr_SHELDUS, here("data/SHELDUS/hurr_SHELDUS_1965-2012_df.csv"))

```


# Import Deryugina

## DATA NOTE

"allYRS_Final_dataset_processed.dta" - this is a dataset prepared in Stata, using the "AllYEARS_Dery_DataProc_justClean.do" do file, which is just the Deryugina cleaning file, modified so it works with my file structure.

In other words,"allYRS_Final_dataset_processed.dta" is the data Deryugina ran her analysis on.

## Import

```{r}

dery_df <- read_dta(here("data/Stata/allYRS_Final_dataset_processed.dta")) %>% 
  # Character version of county_fips
  mutate(FIPS_5 = as.character(county_fips)) %>% 
  mutate(
    FIPS_5 = case_when(
      str_length(FIPS_5) == 5 ~ FIPS_5,
      str_length(FIPS_5) == 4 ~ str_c("0", FIPS_5))
    )

```

# Join

```{r eval=TRUE}

# Function to join datasets and filter
join_and_filter <- function(dery_df, hurr_SHELDUS) {
  dery_df %>%
    left_join(hurr_SHELDUS, by = c("FIPS_5" = "FIPS_5", "year" = "Year")) %>%
    mutate(pre_79 = if_else(year < 1979, 1, 0),
           post_2002 = if_else(year > 2002, 1, 0),
           across(starts_with("PropertyDmg"), ~replace_na(.x, 0))
    )
}

```

```{r eval=TRUE}

# Join the datasets and filter
HrSHLD_dery_df <- join_and_filter(dery_df, hurr_SHELDUS)

```

## Drop Datasets

```{r}

rm(dery_df, hurr_SHELDUS)

```

# Treatment Variables

NOTE: The notes on these variables I wrote for the Hill Final Paper, so they might not all make sense in this context.

I have added 'SHLD' and 'Dery' suffixes to treatment variables, to indicate whether they come from SHELDUS or Deryugina hurricane indicators.

## Hurricane?

These are the different indicators for the county-year when a hurricane hit.

Reminder that, for this code, the SHELDUS indicators will not be used, but the code is still here.

### Deryugina

'hurricane' is the year a county had a hurrican pass through, with winds of more than 74mph.

### SHELDUS

#### Any Hurricane damage

```{r eval=TRUE}

HrSHLD_dery_df <- HrSHLD_dery_df %>%
  # Serious hurricane damage
  mutate(any_hurrDMG_SHLD = if_else(PropertyDmg_Adj2022 > 0 & year >= 1979 & year <= 2002, 1, 0))

```

#### Serious Hurricane damage

NOTE: I am NOT using this serious threshold for treatments, in the hurricane case!! Just leaving the code here, in case.

```{r eval=FALSE}

# Define serious hurricane damage at 50th percentile of SHELDUS counties

serious_threshold <- HrSHLD_dery_df %>%
  filter(SHELDUS == 1 & PropertyDmgPerCapita_Adj2022 > 0) %>%  # Filter rows where SHELDUS == 1 $ drop zeroes
  summarise(quantile_80 = quantile(PropertyDmgPerCapita_Adj2022, 0.5, na.rm = TRUE)) %>%
  pull(quantile_80)  # Extract the quantile value

# Check the calculated threshold
serious_threshold

# Add variable
HrSHLD_dery_df <- HrSHLD_dery_df %>%
  # Serious hurricane damage
  mutate(serious_hurr = if_else(PropertyDmgPerCapita_Adj2022 >= serious_threshold, 1, 0))

```

## 'group' variable

Used for Callaway and Sant'Anna (2021) Group-Time ATT. I also use the 'group' variable to build other treatment variables.

'group' is the first year a county has any hurricane, and is 0 if never-treated.

### SHELDUS

```{r eval=TRUE}

HrSHLD_dery_df <- HrSHLD_dery_df %>%
  # Group by county
  group_by(FIPS_5) %>%
  # Sort by year within each county
  arrange(FIPS_5, year) %>%
  mutate(
    first_hurr_year_SHLD = ifelse(any(any_hurrDMG_SHLD == 1), min(year[any_hurrDMG_SHLD == 1]), NA_real_),
    group_SHLD = ifelse(is.na(first_hurr_year_SHLD), 0, first_hurr_year_SHLD)
  ) %>%
  ungroup()

```

### Deryugina

```{r eval=TRUE}

HrSHLD_dery_df <- HrSHLD_dery_df %>%
  arrange(FIPS_5, year) %>%
  group_by(FIPS_5) %>%
  mutate(
    first_hurr_year_Dery = ifelse(any(hurricane == 1), min(year[hurricane == 1]), NA_real_),
    group_Dery = ifelse(is.na(first_hurr_year_Dery), 0, first_hurr_year_Dery)
  ) %>%
  ungroup()

```

## 'post_treat' variable

(Previously was called 'treatment')

'post_treat' is an indicator for the post-treatment period for each treated county.It is 1 for the year the hurricane happened, and any year after.

### SHELDUS

```{r eval=T}

HrSHLD_dery_df <- HrSHLD_dery_df %>%
  # Group by county
  group_by(FIPS_5) %>%
  # Sort by year within each county
  arrange(FIPS_5, year) %>%
  # Create the treatment variable
  mutate(post_treat_SHLD = cummax(any_hurrDMG_SHLD)) %>%
  ungroup()

```

### Deryugina

```{r eval=TRUE}

HrSHLD_dery_df <- HrSHLD_dery_df %>%
  # Group by county
  group_by(FIPS_5) %>%
  # Sort by year within each county
  arrange(FIPS_5, year) %>%
  # Create the treatment variable
  mutate(post_treat_Dery = cummax(hurricane)) %>%
  ungroup()

```

## 'time_to_treat' & 'ever_treated'

Variables for 'fixest' package.

'time_to_treat' is the event study, with -1000 for never-treated counties (following 'fixest' documentation).

'ever_treated' is a dummy for whether a county is ever treated. (Previously 'treat')

### SHELDUS

```{r eval=T}

HrSHLD_dery_df <- HrSHLD_dery_df %>% 
  mutate(
    time_to_treat_SHLD = ifelse(group_SHLD > 0, year - group_SHLD, -1000),
    ever_treated_SHLD = ifelse(group_SHLD > 0, 1, 0)
  )

```

### Deryugina

```{r eval=TRUE}

HrSHLD_dery_df <- HrSHLD_dery_df %>% 
  mutate(
    time_to_treat_Dery = ifelse(group_Dery > 0, year - group_Dery, -1000),
    ever_treated_Dery = ifelse(group_Dery > 0, 1, 0)
  )

```

## 'treat_DiD'

'treat_DiD' is an interaction of 'post_treat' (the post-treatment period indicator) and 'ever_treated' (the ever-treated indicator). The coefficient of this interaction is the traditional difference-in-differences ATT estimate.

### SHELDUS

```{r eval=T}

HrSHLD_dery_df <- HrSHLD_dery_df %>% 
  mutate(
    treatment_post_SHLD = post_treat_SHLD * ever_treated_SHLD
  )

```

### Deryugina

```{r eval=TRUE}

HrSHLD_dery_df <- HrSHLD_dery_df %>% 
  mutate(
    treatment_post_Dery = post_treat_Dery * ever_treated_Dery
  )

```

# Other Data Management

## Save .rds

```{r}

saveRDS(HrSHLD_dery_df, here("data/SHELDUS/SHELDUS_Deryugina.rds"))

```


## Export Full Data

```{r}

# Full Data
write_csv(HrSHLD_dery_df, here("data/SHELDUS/SHELDUS_Deryugina.csv"))

```

### Sample Hurr State (note)

The 'hs_' file, for the states in the 'sample_hurr_state' group, is not included here. That is because I used the Deryugina data set AFTER her DataProcessingAnalysis.do file, and this file filters for only the Sample Hurr States.

# Sample

## Simple

```{r eval=FALSE}

sample_df <- HrSHLD_dery_df %>%
  slice_sample(n = 100)

```

```{r eval=FALSE}

write_csv(sample_df, here("data/SHELDUS/sample_dery_df.csv"))

```

## Sample Counties

```{r eval=FALSE}

# Sample 10% of the unique counties
sampled_counties <- HrSHLD_dery_df %>%
  distinct(FIPS_5) %>%
  slice_sample(prop = 0.10) %>%
  pull(FIPS_5)

# Filter your dataset to include only the data from the sampled counties.
smpl_counties_df <- HrSHLD_dery_df %>%
  filter(FIPS_5 %in% sampled_counties)

```

### Treated Counties

```{r eval=FALSE}

# Sample 5% of the unique counties
sampled_Tcounties <- HrSHLD_dery_df %>%
  filter(ever_treated_SHLD == 1) %>% 
  distinct(FIPS_5) %>%
  slice_sample(prop = 0.05) %>%
  pull(FIPS_5)

# Filter your dataset to include only the data from the sampled counties.
smpl_Tcounties_df <- HrSHLD_dery_df %>%
  filter(FIPS_5 %in% sampled_Tcounties) %>% 
  # Check SHELDUS Cleaning
  select(year | FIPS_5:treatment_post_SHLD)

```



